Multilevel generalizations of collective spin dynamics and squeezing

### 2019-10-26 collective_sun/multilevel_spin_notes.tex
write the excited-state projectors in second-order perturbation theory using eigenspace projectors, rather than eigenstates

Here we derive the effective Hamiltonian induced on the fully
symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC_B} in the weak SOC limit
$\abs{\phi}\ll1$.  For brevity, we work in units with the lattice
spacing $a=1$ throughout this section.  Perturbations $V$ to the
interaction Hamiltonian $H_{\t{int}}$ induce the following effective
Hamiltonians on the fully symmetric manifold
$\M_\FS$\cite{bravyi2011schrieffer, perlin2019effective}:
\begin{align}
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
  &&
  H_{\t{eff}}^{(2)} = -\P_0 V \E V \P_0,
  &&
  \P_0 = \sum_{E_{\varepsilon}=0} \op{E_{\varepsilon}},
  &&
  \E \equiv \sum_{E_\varepsilon>0} \f{\op{E_\varepsilon}}{E_\varepsilon},
  \label{eq:pert}
\end{align}
where $\ket{E_\varepsilon}$ is an eigenstate of the interaction
Hamiltonian $H_{\t{int}}$ with energy $E_\varepsilon$ relative to the
interaction energy of fully symmetric states
$\ket{\psi_\FS}\in\M_\FS$.  The sum over all $E_\varepsilon>0$ is a
sum over a complete basis of states $\set{\ket{E_\varepsilon}}$ for
the orthogonal complement of $\M_\FS$.

@@

In order to expand the second-order effective Hamiltonian
$H_{\t{eff}}^{(2)}$ induced on the fully symmetric manifold $\M_\FS$
by the SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
\eqref{eq:H_lat_SOC_A} when the SOC angle $\abs{\phi}\ll1$, we note
that $w_L\p{q}$ in \eqref{eq:H_lat_SOC_A} takes the form
\begin{align}
  w_L\p{q} = \f12 \sum_{s\in\set{\pm1}} \chi_{Ls} e^{isq},
  \label{eq:w_chi}
\end{align}
for some appropriately defined coefficients $\chi_{Ls}$, which implies
that
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -\f{t}{2} \sum_{\substack{L\\s\in\set{\pm1}}}
  \chi_{Ls} A_L^{(\phi)} \Z_{Ls},
  &&
  \Z_{Ks} \equiv \sum_q e^{isq} D_{L,0}^{(q)},
\end{align}
and in turn
\begin{align}
  H_{\t{eff}}^{(2)}
  = -\f{t^2}{4} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
  \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
  \label{eq:H_eff_2_start}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold
$\M_\FS$, and $\E$ is a sum over excited-state projectors with
energetic suppression factors, defined in \eqref{eq:pert}.  The
remainder of this section is devoted to the task of simplifying
\eqref{eq:H_eff_2_start} to arrive at our final result in
\eqref{eq:H_eff_2}, expressed in purely in terms of collective drive
operators.  We will break up this task into three steps:
\begin{enumerate*}
\item eliminating the excited-state projectors in $\E$,
\item evaluating the product of operators in \eqref{eq:H_eff_2_start},
  and
\item evaluating the sum over $r,s\in\set{\pm1}$.
\end{enumerate*}

@@

these spin wave operators that are off-diagonal with respect to the

@@

  \EE_k\sp{X\p{k}} \equiv \f1N \sum_k X\p{k},

>>>>

Here we derive the effective Hamiltonian induced on the fully
symmetric manifold $\M_\FS$ by the single-particle SOC Hamiltonian
$H_{\t{lat}}^{(\phi)}$ in \eqref{eq:H_lat_SOC_B} in the weak SOC limit
$\abs{\phi}\ll1$.  For brevity, we work in units with the lattice
spacing $a=1$ throughout this section.  Perturbations $V$ to the
interaction Hamiltonian $H_{\t{int}}$ induce the following effective
Hamiltonians on the fully symmetric manifold
$\M_\FS$\cite{bravyi2011schrieffer, perlin2019effective}:
\begin{align}
  H_{\t{eff}}^{(1)} = \P_0 V \P_0,
  &&
  H_{\t{eff}}^{(2)} = -\P_0 V \E V \P_0,
  &&
  \E \equiv \sum_{\Delta>0} \f{\P_\Delta}{\Delta},
  \label{eq:pert}
\end{align}
where $\P_\Delta$ is a projector onto the eigenspace of the
interaction Hamiltonian $H_{\t{int}}$ corresponding to an energy
$\Delta$ above that of fully symmetric states
$\ket{\psi_\FS}\in\M_\FS$.

@@

In order to expand the second-order effective Hamiltonian
$H_{\t{eff}}^{(2)}$ induced on the fully symmetric manifold $\M_\FS$
by the SOC Hamiltonian $H_{\t{lat}}^{(\phi)}$ in
\eqref{eq:H_lat_SOC_A} when the SOC angle $\abs{\phi}\ll1$, we note
that $w_L\p{q}$ in \eqref{eq:H_lat_SOC_A} takes the form
\begin{align}
  w_L\p{q} = \f12 \sum_{s\in\set{\pm1}} \chi_{Ls} e^{isq},
  \label{eq:w_chi}
\end{align}
for some appropriately defined coefficients $\chi_{Ls}$, which implies
that
\begin{align}
  H_{\t{lat}}^{(\phi)}
  = -\f{t}{2} \sum_{\substack{L\\s\in\set{\pm1}}}
  \chi_{Ls} A_L^{(\phi)} \Z_{Ls},
  &&
  \Z_{Ks} \equiv \sum_q e^{isq} D_{L,0}^{(q)},
\end{align}
and in turn
\begin{align}
  H_{\t{eff}}^{(2)}
  = -\f{t^2}{4} \sum_{\substack{J,K\\r,s\in\set{\pm1}}}
  \chi_{Jr} \chi_{Ks} A_J^{(\phi)} A_K^{(\phi)}
  \P_0 \Z_{Jr} \E \Z_{Ks} \P_0,
  \label{eq:H_eff_2_start}
\end{align}
where $\P_0$ is a projector onto the fully symmetric manifold
$\M_\FS$, and $\E$ is a sum over projectors onto excited-state
manifolds with corresponding energetic suppression factors, defined in
\eqref{eq:pert}.  The remainder of this section is devoted to the task
of simplifying \eqref{eq:H_eff_2_start} to arrive at our final result
in \eqref{eq:H_eff_2}, expressed in purely in terms of collective
drive operators that are implicitly restricted to the fully symmetric
manifold $\M_\FS$.  We will break up our task into three steps:
\begin{enumerate*}
\item eliminating the excited-state projectors in $\E$,
\item evaluating the product of operators in \eqref{eq:H_eff_2_start},
  and
\item evaluating the sum over $r,s\in\set{\pm1}$.
\end{enumerate*}

@@

these spin-wave operators that are off-diagonal with respect to the

@@

  \EE_k\sp{X_k} \equiv \f1N \sum_k X_k,
